---
title: "Machine Learning Course Project"
author: "Andrew Wilson"
date: "Tuesday, December 22, 2015"
output: html_document
---

#Summary
The intent of this report is to provide a reproducible method of predicting the performance quality of different exercises based on data collected by accelerometers.  The analysis will be conducted using cross validation techniques to determine a machine learning algorithm with which to build a prediction model.

#Getting Started
The analyis in this study will rely on certain packages that will need to be installed and loaded into R.
```{r}
library(caret)
```
To ensure consistent results, set the following seed in R:
```{r}
set.seed(1334)
```
Both training and test sets for this analysis can be accessed and loaded into memory using the following URLs:
```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```
Training dataset is further separated out into a training subset for building the model, and a testing subset.

```{r}
inTrain<-createDataPartition(training$classe,p=.6, list=FALSE)
subTrain<-training[inTrain,]
subTest<-training[-inTrain,]
```

#Exploratory Analysis and Data Cleaning
This section is mainly concerned with removing unecessary variables that will interfere with the prediction model.  First variables with near zero variance are detected and removed:
```{r}
nearZero<-nearZeroVar(subTrain,saveMetrics =TRUE)
nearZeroVars<-row.names(nearZero[which(nearZero$nzv=='TRUE'),])
remove<-names(subTrain) %in% nearZeroVars
subTrain<-subTrain[!remove]
```

The data has some sparse fields that are heavily populated with NA values.  These will be removed to try and cut down on noise in the prediction.

```{r}
remove2<-colSums(is.na(subTrain)) > 5000
subTrain<-subTrain[!remove2]
```
 
To further simplify the prediction model,  all columns not related to acceleremoter data will be removed.
```{r}
remove3<-c("X","user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp","num_window")
remove3<-names(subTrain) %in% remove3
subTrain<-subTrain[!remove3]
```

The same subTest and testing datasets will then need to be processed with the same columns removed. It is important to note that the testing dataset does not include the output "classe" variable so it will need to be removed from the column names used to subset this dataset.

```{r}
colName<-colnames(subTrain)
colName2<-colnames(subTrain[,-53])
subTest<-subTest[colName]
testing<-testing[colName2]
```

##Random Forest Prediction Algorithm
To predict the "classe" variable I will use a Random Forest algorithm.




