---
title: "Machine Learning Course Project"
author: "Andrew Wilson"
date: "Tuesday, December 22, 2015"
output: html_document
---

#Summary
The intent of this report is to provide a reproducible method of predicting the performance quality of different exercises based on data collected by accelerometers.  The analysis will be conducted using cross validation techniques to determine a machine learning algorithm with which to build a prediction model.

#Getting Started
The analyis in this study will rely on certain packages that will need to be installed and loaded into R.
```{r}
library(caret)
```
To ensure consistent results, set the following seed in R:
```{r}
set.seed(1334)
```
Both training and test sets for this analysis can be accessed and loaded into memory using the following URLs:
```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```
Training dataset is further separated out into a training subset for building the model, and a testing subset.

```{r}
inTrain<-createDataPartition(training$classe,p=.6, list=FALSE)
subTrain<-training[inTrain,]
subTest<-training[-inTrain,]
```

